\Large {\bf Protocol-Guided Analysis of Post-silicon Traces Under Limited Observability}

### 2015-09-28 paper.tex
added one more experiment into case study
Section III, updated algorithm 1, added a small example to illustrate
the trace analysis.

section V: results from one more experiment are added.

Case study I is removed to save space.

A key component of post-silicon validation of SoC designs is
to correlate trace from silicon execution with the intended
system-level transactions.  An SoC design is typically
composed of a large number of pre-designed hardware or
software blocks (often referred to as ``intellectual
properties'' or ``IPs'') that coordinate through complex
protocols to implement the system-level behavior.  Any
execution trace of the system involves a large number of
interleaved instances of these protocols.  For example,
consider a smartphone executing a usage scenerio where the
end-user browses the Web while listening to music and
sending and receiving occasional text messages.  Typical
post-silicon validation use-case involve exercising such
scenarios.  An execution trace would involve activities from
the CPU, audio controller, display controller, wireless
radio antenna, etc., reflecting the interleaved execution of
several communication protocols.  On the other hand, due to
observability limitations, only a small small number of
participating signals can be actually traced during silicon
execution.  Furthermore, due to electrical perturbations,
silicon data can be noisy, lossy, and ambiguous.
Consequently, it is non-trivial to identify all
participating protocols and pinpoint the interleaving that
results in an observed trace.

In this paper, we consider the problem of reconstructing
protocol-level behavior from silicon traces in SoC designs.
Given a collection of system-level communication protocols
and a trace of (partially observed) hardware signals, our
approach infers, with a certain measure of confidence, the
protocol instances (and their interleavings) being exercised
by the trace.  The approach is based on a formalization of
system-level transactions via labeled Petri-Nets, which are
capable of describing sequencing, concurrency, and choices
over system events.  We develop algorithms to infer
system-level transactions from traces with missing, noisy,
and ambiguous signal values, together with an estimate of
confidence on the inference.  We demonstrate our approach on
two SoC case studies: a pedagogical SystemC prototype and a
more realistic implementation constructed within the GEM5
environment~\cite{Binkert2011}.

\begin{figure*}[tb]
\begin{center}
\begin{tabular}{cc}
\begin{minipage}{3in}
\includegraphics[height=2in]{figures/bpmn-flow-ex}
\end{minipage}
&
\begin{minipage}{3in}
\resizebox{2.8in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(t1)	at	($(p1) + (0,-1.1)$) {$t_1:\langle {\tt Driver:Device:Load} \_fw\rangle$};
	\node[circ] 	(p2) 		at ($(t1) + (0,-1.1)$) {$p_2$};
	\node[rectbox]	(t2)	at	($(p2) + (0,-1.1)$) {$t_2:\langle {\tt Device:CE:Auth\_req} \rangle$};
	\node[circ] 	(p3) 		at ($(t2) + (0,-1.1)$) {$p_3$};
	\node[rectbox]	(t3)	at	($(p3) + (0,-1.1)$) {$t_3:\langle {\tt CE:Device:Auth\_resp}\rangle$};
	\node[circ] 	(p4) 		at ($(t3) + (0,-1.1)$) {$p_4$};
	\node[rectbox]	(t4)	at	($(p4) + (-3,-1.1)$) {$t_4:\langle {\tt Device:Driver:Report} \rangle$};
	\node[rectbox]	(t5)	at	($(p4) + (3,-1.1)$) {$t_5:\langle {\tt Device:CE:Ack} \rangle$};
	\node[circ,fill=black] 	(term1) 		at ($(t4) + (0,-1.1)$) {$~~~$};
	\node[circ,fill=black] 	(term2) 		at ($(t5) + (0,-1.1)$) {$~~~$};

@@

An SoC design involves integration of a number of IPs that
communicate through complex protocols.  Such system-level
protocols are typically specified in architecture documents
as message flow diagrams.  As illustration,
Fig~\ref{flow-spec-ex}(a) shows one such diagram for a
protcol to authenticate and load a firmware during system
boot for firmware upgrade.  During validation, the system
under debug (SUD) exercises some complex system-level
use-case which involves interleaved execution of possibly a
large number of such protocols.  A trace of a small number
of hardware signals is then shipped off-chip analysis.  The
off-chip analysis includes two broad phases: (1)~trace
abstraction, and (2)~trace interpretation.  Trace
abstraction maps the hardware trace into higher-level
architectural constructs, \eg, messages, operations, etc.: a
message such as {\tt Authorization request} may be
implemented in hardware through a Boolean or temporal
combination of specific hardware signals in the NoC fabric
between {\tt Device} and {\tt CE}, \eg, as a sequence
containing a header, a specific value of a sequence of data
words, etc.  We will refer to such architectural constructs
as {\em protocol events} or {\em flow events}.  Note that
due to limited observability, it may not be possible to map
a given set of (observed) hardware signals to uniquely to a
flow event.  Finally, the trace may include several
instances of the same protocol executing concurrently, \eg,
a firmware authentication protocol may be invoked when
another instance of the protocol has not completed.

Trace interpretation entails mapping flow events created
during trace abstraction to system-level protocols in order
to identify the set of protocol instances (and
interleavings) responsible for creating the observed
behavior.  The trace may identify a problem in the protocols
themselves, \eg an interleaving of some protocol executions
may lead to an unexpected message being sent or cause the
system to crash.  More commonly, one finds a bug in the {\em
  implementation} of the protocol, \ie, a trace inconsistent
with any possible interleaving of the protocol executions.
Identifying these problems involves significant human
expertise, and can often take days to weeks of effort.

@@

\noindent 
{\bf Notations and formalization.}  We assume that the set
of system flows is provided as a collection ${\vec{F}}$ of
LPNs.  We define a {\em flow execution scenario} to be the
set of pairs $(F_{i,j}, s_{i,j})$ where $F_{i,j}$ is the
$j$th instance of flow $F_i \in {\vec{F}}$, and $s_{i,j}$ is
a state of $F_{i,j}$.  A flow execution scenario indicates
the set of protocols and the number of instances of a
particular protocol are activated and their corresponding
current states.  Since we assume full observability, we view
an {\em observed trace} $\rho = e_1e_2\ldots e_n$ is a
sequence of events.  Given an observed trace $\rho$, the
goal of of trace interpretation is to construct a set of
candidate flow execution scenarios whose execution can
create the sequence of events in $\rho$.  We call such
execution scenarios {\em compliant} with $\rho$.  Let
$\mathit{accept(F_{i,j}, s_{i,j}, e)}$ be a function that
determines if event $e$ can be emitted by $F_{i,j}$ in state
$s_{i,j}$.  Formally, $\mathit{accept(F_{i,j}, s_{i,j}, e)}$
returns $(F_{i,j}, s'_{i,j})$ if there exists a transition
$t$ in $F_i$ such that $L(t) = e$ and $\bullet t \subseteq
s_{i,j}$.  In this case, $s'_{i,j} = (s_{i,j} - \bullet t)
\cup t\bullet$.  It returns $\emptyset$ if no such $t$
exists in $F_i$.

@@

Given an obsered trace $\rho$ and the set $\vec{F}$ of LPNs,
Algorithm~\ref{algo:compliance} provides the basic steps for
computing a set of compliant flow execution scenarios.
\begin{algorithm}[ht]
\DontPrintSemicolon
$\mathit{Scen} \gets \emptyset$\;
\ForEach {$h, \; 1 \leq h \leq n $} {
$\mathit{found} \gets {\tt true}$ \;
  \ForEach {$(F_{i,j}, s_{i,j}) \in \mathit{scen}$} {
    \If {$\mathit{accept}(F_{i,j}, s_{i,j}, e_h) = (F_{i,j}, s'_{i,j})$} {
      $\mathit{scen} \gets (\mathit{scen} - (F_{i,j}, s_{i,j})) \cup (F_{i,j}, s'_{i,j})$\;
      $\mathit{scen} \gets \mathit{scen} \cup \mathit{Scen}'$\;
      $\mathit{found} \gets {\tt false}$ \;
    }
  }
  \ForEach {$F_i \in \vec{F}$} {
      create a new instance $F_{i, j+1}$ \;
      \If {$\mathit{accept}(F_{i,j+1},\mathit{init}_{i,j+1}, e_h) = (F_{i,j+1}, s'_{i,j+1})$} {
    $\mathit{scen} \gets \mathit{scen} \cup (F_{i,j+1}, s'_{i,j+1})$ \;
        $\mathit{found} \gets {\tt false}$ \;
      }
    }
  \If {$\mathit{found} == {\tt true}$} {
    \Return ${\tt Inconsistent}$\;
  }
}
\Return $\mathit{scen}$ \;
\caption{$\textsc{Check-Compliance}(\vec{F}, \, \rho)$}
\label{algo:compliance}
\end{algorithm}
The algorithm operates by keeping track (in variable
$\mathit{scen}$) of a set of candidate protocol instances
compliant with each prefix of $\rho$.  At each iteration, we
update this set by either extending some member of
$\mathit{scen}$ or initiating a new protocol instance.  If
neither is possible then we report that the trace is {\tt
  inconsistent}, \ie, there is no possible interleaving of
the protocol instances from $\vec{F}$ that is compliant with
$\rho$.

%% Given a trace of flow events $\rho = e_1e_2\ldots e_n$, the
%% trace interpretation algorithm starts with an empty set of
%% of flow execution scenario $\mathit{Scen} = \emptyset$.
%% Then, for each $e_h$ where $1 \leq h \leq n$ starting $h=1$,
%% and for each $\mathit{scen} \in \mathit{Scen}$, the
%% following two steps are performed.
%% \begin{description}
%% \item[Step 1]~~For each $(F_{i,j}, s_{i,j}) \in
%%   \mathit{scen}$, if $\mathit{accept}(F_{i,j}, s_{i,j}, e_h)
%%   = (F_{i,j}, s'_{i,j})$, create a new scenario
%%   $\mathit{scen}' = (\mathit{scen} - (F_{i,j}, s_{i,j}))
%%   \cup (F_{i,j}, s'_{i,j})$, which is added into
%%   $\mathit{Scen}'$.

@@

%=====
\section{Case Study}

The proof concept of the proposed trace analysis method is demonstrated on a transaction level model of a simple SoC design with two CPU cores, memory, and a peripheral device connected by an interconnect as shown in Fig.~\ref{fig-SoC-ex}.  Since the analysis method presented in this paper is communication centric, the detailed computations of these blocks are not modeled.  Instead, the modeling is focused on how they participate in flows for different system level use cases.

\begin{figure}[tb]
\begin{center}
\includegraphics[width=2in]{figures/simple-ex}
\caption{A simple SoC example.}
\label{fig-SoC-ex}
\end{center}
\end{figure}

\subsection{Flow Specification}

In this case study, four system flows are implemented in the simple SoC model. They include cache coherent memory access operations and a memory-mapped peripheral read operation initiated from the CPUs, a message signaled interrupt operation initiated from the peripheral device.  These flow specifications capture how messages are exchanged for different use cases.  In this model, a message is defined with the following format, $(\mbox{Src}, \mbox{Dest}, \mathit{Cmd}, \mathit{Addr})$, where Src and Dest refer to the source and destination components of messages, $\mathit{Cmd}$ refers to the operations that the destination component should perform, and $\mathit{Addr}$ refers to memory addresses where $\mathit{Cmd}$ applies.  $\mathit{Cmd}$ can be memory accesses or not.  If it is not, then the $\mathit{Addr}$ field of messages is ignored.  In this case study, memory mapped IO mechanism is used.  Furthermore, detailed memory addresses are not modeled. Instead, the address space is partitioned to main memory addresses and peripheral addresses.  In messages, the $\mathit{Addr}$ field is replaced with either $M$ representing a memory address or $P$ representing an address to a peripheral device.

The LPN as shown in Fig.~\ref{fig-cpu-write} specifies a system flow where {CPU2} initiates a memory write operation.   In this flow, CPU2 initiates a memory write request followed by a data valid message to the interconnect.  The data valid messages are used to model availability or validity of data for transfer.  Concurrently, the interconnect inquiries CPU1 if it holds a more updated version of the data by sending a memory write message $\mathit{msg_2}$.   CPU1 generates one of two possible responses.  If CPU1 holds the more updated data in its cache in the same memory space that CPU2 intends to write, a cache hit message $\mathit{msg_4}$ followed by $\mathit{msg_6}$ are sent to the interconnect.  Otherwise, CPU1 sends a cache miss message $\mathit{msg_5}$.  After getting the response from CPU1, Interconnect sends a write request to the memory unit.  This flow is symmetric for CPU1.

\begin{figure}[tb]
\begin{center}
\resizebox{2.8in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(msg1)	at	($(p1) + (0,-1.1)$) {$msg_1$};
	\node[circ] 	(p2) 		at ($(msg1) + (-3,0)$) {$p_2$};
	\node[circ] 	(p3) 		at ($(msg1) + (3,0)$) {$p_3$};
	\node[rectbox]	(msg2)	at	($(p2) + (0,-1.1)$) {$msg_2$};
	\node[rectbox]	(msg3)	at	($(p3) + (0,-1.5)$) {$msg_3$};
	\node[circ] 	(p4) 		at ($(msg2) + (0,-1.1)$) {$p_4$};
	\node[rectbox]	(msg4)	at	($(p4) + (0,-1.1)$) {$msg_4$};
	\node[rectbox]	(msg5)	at	($(p4) + (3,0)$) {$msg_5$};
	\node[circ] 	(p6) 		at ($(msg4) + (0,-1.1)$) {$p_6$};
	\node[circ] 	(p5) 		at ($(msg3) + (0,-1.4)$) {$p_5$};
	\node[rectbox,rotate=90]	(msg6)	at	($(p6) + (1.5,0)$) {$msg_6$};
	\node[circ] 	(p7) 		at ($(msg6) + (1.5,0)$) {$p_7$};
	\node[rectbox]	(msg7)	at	($(p5) + (0,-1.5)$) {$msg_7$};
	\node[terminal]	(p8)	at	($(msg7) + (2,0)$) {$p_8$};
	% Edges
	\draw[->,weight3]	(p1) -- (msg1);
	\draw[->,weight3]	(msg1) -- (p2);
	\draw[->,weight3]	(msg1) -- (p3);
	\draw[->,weight3]	(p2) -- (msg2);
	\draw[->,weight3]	(msg2) -- (p4);
	\draw[->,weight3]	(p4) -- (msg4);
	\draw[->,weight3]	(p4) -- (msg5);
	\draw[->,weight3]	(msg4) -- (p6);
	\draw[->,weight3]	(p6) -- (msg6);
	\draw[->,weight3]	(msg6) -- (p7);
	\draw[->,weight3]	(p7) -- (msg7);
	\draw[->,weight3]	(msg7) -- (p8);
	\draw[->,weight3]	(msg5) -- (p7);
	\draw[->,weight3]	(p3) -- (msg3);
	\draw[->,weight3]	(msg3) -- (p5);
	\draw[->,weight3]	(p5) -- (msg7);

	\node[]	(msg-def)	at	($(p1) + (0,-8.5)$) {
	\begin{minipage}{3in}
	Definition of the messages:
\[
\begin{array}{clllll}
msg_1: & (\mbox{CPU2},	& \mbox{Icon},		& \mathit{Wr}, 	& M) \\
msg_2: & (\mbox{Icon},	& \mbox{CPU1}, 	& \mathit{Wr},			& M) \\
msg_3: & (\mbox{CPU2},	& \mbox{Icon}, 	& \mathit{DVal},	& -) \\
msg_4: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Hit},	& -) \\
msg_5: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Miss},	& -) \\
msg_6: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{DVal},	& -) \\
msg_7: & (\mbox{Icon},	& \mbox{Mem}, 		& \mathit{Wr},			& M) \\
\end{array}
\]
\end{minipage}
		};
\end{tikzpicture}
}
\caption{Flow specification ($F_1$) of a cache coherent write operation initiated from \texttt{CPU2}.}
\label{fig-cpu-write}
\end{center}
\end{figure}

The LPN specification as shown in Fig.~\ref{fig-cpu-read} captures the system flow where CPU2 initiates a memory read operation.  Basically, CPU2 sends a memory read message to Interconnect, which then generates two concurrent threads, one checks if CPU1 has the more updated data in the memory space for the read operation, and the other thread to get data from the memory.  Once Interconnect gets both responses from CPU1 and memory, it synchronizes the responses, and writes the correct data to the CPU2's cache and memory in parallel.  Again, this specification is symmetric for CPU1.

\begin{figure}[htbp]
\begin{center}
\resizebox{3.2in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(msg1)	at	($(p1) + (0,-1.1)$) {$msg_1$};
	\node[circ] 	(p2) 		at ($(msg1) + (-3,0)$) {$p_2$};
	\node[circ] 	(p3) 		at ($(msg1) + (2,0)$) {$p_3$};
	\node[rectbox]	(msg2)	at	($(p2) + (0,-1.1)$) {$msg_2$};
	\node[rectbox]	(msg3)	at	($(p3) + (2,0)$) {$msg_3$};
	\node[circ] 	(p4) 		at ($(msg2) + (0,-1.1)$) {$p_4$};
	\node[rectbox]	(msg4)	at	($(p4) + (0,-1.1)$) {$msg_4$};
	\node[rectbox]	(msg5)	at	($(p4) + (3,0)$) {$msg_5$};
	\node[circ] 	(p6) 		at ($(msg4) + (0,-1.1)$) {$p_6$};
	\node[circ] 	(p5) 		at ($(msg3) + (0,-1.1)$) {$p_5$};
	\node[rectbox,rotate=90]	(msg6)	at	($(p6) + (1.5,0)$) {$msg_6$};
	\node[circ] 	(p7) 		at ($(msg6) + (1.5,0)$) {$p_7$};
	\node[rectbox]	(msg7)	at	($(p5) + (0,-1.1)$) {$msg_7$};
	\node[circ]	(p8)	at	($(msg7) + (0,-1.1)$) {$p_8$};
	\node[rectbox]	(msg8)	at	($(p8) + (0,-1.1)$) {$msg_8$};
	\node[circ]	(p9)	at	($(msg8) + (-1.5,-1.1)$) {$p_9$};
	\node[circ]	(p10)	at	($(msg8) + (1.5,-1.1)$) {$p_{10}$};
	\node[rectbox]	(msg9)	at	($(p9) + (0,-1.1)$) {$msg_9$};
	\node[rectbox]	(msg10)	at	($(p10) + (0,-1.1)$) {$msg_{10}$};
	\node[terminal]	(p11)	at	($(msg9) + (0,-1.1)$) {$p_{11}$};
	\node[terminal]	(p12)	at	($(msg10) + (0,-1.1)$) {$p_{12}$};

	% Edges
	\draw[->,weight3]	(p1) -- (msg1);
	\draw[->,weight3]	(msg1) -- (p2);
	\draw[->,weight3]	(msg1) -- (p3);
	\draw[->,weight3]	(p2) -- (msg2);
	\draw[->,weight3]	(msg2) -- (p4);
	\draw[->,weight3]	(p4) -- (msg4);
	\draw[->,weight3]	(p4) -- (msg5);
	\draw[->,weight3]	(msg4) -- (p6);
	\draw[->,weight3]	(p6) -- (msg6);
	\draw[->,weight3]	(msg6) -- (p7);
	\draw[->,weight3]	(p7) -- (msg8);
	\draw[->,weight3]	(msg7) -- (p8);
	\draw[->,weight3]	(msg5) -- (p7);
	\draw[->,weight3]	(p3) -- (msg3);
	\draw[->,weight3]	(msg3) -- (p5);
	\draw[->,weight3]	(p5) -- (msg7);
	\draw[->,weight3]	(p8) -- (msg8);
	\draw[->,weight3]	(msg8) -- (p9);
	\draw[->,weight3]	(msg8) -- (p10);
	\draw[->,weight3]	(p9)  -- (msg9);
	\draw[->,weight3]	(p10)  -- (msg10);
	\draw[->,weight3]	(msg9) -- (p11);
	\draw[->,weight3]	(msg10) -- (p12);

	\node[]	(msg-def)	at	($(p1) + (-2.7,-10)$) {
	\begin{minipage}{2in}
	Definition of the messages:
\[
\begin{array}{clllll}
msg_1: & (\mbox{CPU2},	& \mbox{Icon},		& \mathit{Rd}, 		& M) \\
msg_2: & (\mbox{Icon},	& \mbox{CPU1}, 	& \mathit{Rd},	& M) \\
msg_3: & (\mbox{Icon},	& \mbox{Mem}, 		& \mathit{Rd},			& M) \\
msg_4: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Hit},	& -) \\
msg_5: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Miss},& -) \\
msg_6: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{DVal},	& -) \\
msg_7: & (\mbox{Mem},		& \mbox{Icon}, & \mathit{DVal},			& -) \\
msg_8: & (-,		& -, 		& -,				& -) \\
msg_9: & (\mbox{Icon},	& \mbox{Mem}, 		& \mathit{Wr},			& M) \\
msg_{10}: & (\mbox{Icon},	& \mbox{CPU2}, 		& \mathit{Wr},			& M) \\
\end{array}
\]
\end{minipage}
		};

\end{tikzpicture}
}
\caption{Flow specification ($F_2$) of a cache coherent read operation from CPU2.}
\label{fig-cpu-read}
\end{center}
\end{figure}

The LPN specification as shown in Fig.~\ref{fig-peri-read} captures a system flow for CPU initiated memory mapped peripheral read operations.  When CPU2 tries to read the peripheral device (device hereafter), a read message with address $P$ is sent to Interconnect, which then sends this message to CPU1 and the device simultaneously.  When CPU1 sees this message, it responds with a cache miss message as the address $P$ points to the device.  At the meantime, the device responds with a data valid message indicating the availability of the requested data.  Finally, Interconnect synchronizes the both responses, and sends a data valid message back to CPU2.   This specification is also symmetric for CPU1.

>>>>

A key component of post-silicon validation of SoC designs is
to correlate traces from silicon execution with the intended
system-level transactions.  An SoC design is typically
composed of a large number of pre-designed hardware or
software blocks (often referred to as ``intellectual
properties'' or ``IPs'') that coordinate through complex
protocols to implement the system-level behavior.  Any
execution trace of the system involves a large number of
interleaved instances of these protocols.  For example,
consider a smartphone executing a usage scenario where the
end-user browses the Web while listening to music and
sending and receiving occasional text messages.  Typical
post-silicon validation use-cases involve exercising such
scenarios.  An execution trace would involve activities from
the CPU, audio controller, display controller, wireless
radio antenna, etc., reflecting the interleaved execution of
several communication protocols.  On the other hand, due to
observability limitations, only a small number of
participating signals can be actually traced during silicon
execution.  Furthermore, due to electrical perturbations,
silicon data can be noisy, lossy, and ambiguous.
Consequently, it is non-trivial to identify all
participating protocols and pinpoint the interleaving that
results in an observed trace.

In this paper, we consider the problem of reconstructing
protocol-level behavior from silicon traces in SoC designs.
Given a collection of system-level communication protocols
and a trace of (partially observed) hardware signals, our
approach infers, with a certain measure of confidence, the
protocol instances (and their interleavings) being exercised
by the trace.  The approach is based on a formalization of
system-level transactions via labeled Petri-Nets, which are
capable of describing sequencing, concurrency, and choices
over system events.  We develop algorithms to infer
system-level transactions from traces with missing, noisy,
and ambiguous signal values.
% together with an estimate of confidence on the inference.  
We demonstrate our approach on
two SoC case studies: a pedagogical SystemC prototype and a
more realistic implementation constructed within the GEM5
environment~\cite{Binkert2011}.

\begin{figure*}[tb]
\begin{center}
\begin{tabular}{cc}
\begin{minipage}{3in}
\includegraphics[height=2in]{figures/bpmn-flow-ex}
\end{minipage}
&
\begin{minipage}{3in}
\resizebox{2.8in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(t1)	at	($(p1) + (0,-1.1)$) {$t_1:\langle {\tt Driver:Device:Load} \_fw\rangle$};
	\node[circ] 	(p2) 		at ($(t1) + (0,-1.1)$) {$p_2$};
	\node[rectbox]	(t2)	at	($(p2) + (0,-1.1)$) {$t_2:\langle {\tt Device:CE:Auth\_req} \rangle$};
	\node[circ] 	(p3) 		at ($(t2) + (0,-1.1)$) {$p_3$};
	\node[rectbox]	(t3)	at	($(p3) + (0,-1.1)$) {$t_3:\langle {\tt CE:Device:Auth\_resp}\rangle$};
	\node[circ] 	(p4) 		at ($(t3) + (0,-1.1)$) {$p_4$};
	\node[rectbox]	(t4)	at	($(p4) + (-3,-1.1)$) {$t_4:\langle {\tt Device:Driver:Report} \rangle$};
	\node[rectbox]	(t5)	at	($(p4) + (3,-1.1)$) {$t_5:\langle {\tt Device:CE:Ack} \rangle$};
	\node[circ,fill=black] 	(term1) 		at ($(t4) + (0,-1.1)$) {\textcolor{white}{$p_5$}};
	\node[circ,fill=black] 	(term2) 		at ($(t5) + (0,-1.1)$) {\textcolor{white}{$p_6$}};

@@

An SoC design involves integration of a number of IPs that
communicate through complex protocols.  Such system-level
protocols are typically specified in architecture documents
as message flow diagrams.  As illustration,
Fig~\ref{flow-spec-ex}(a) shows one such diagram for a
protcol to authenticate and load a firmware during system
boot for firmware upgrade.  During validation, the system
under debug (SUD) exercises some complex system-level
use-case which involves interleaved execution of possibly a
large number of such protocols.  A trace of a small number
of hardware signals is then shipped off-chip analysis.  The
off-chip analysis includes two broad phases: (1)~trace
abstraction, and (2)~trace interpretation.  Trace
abstraction maps the hardware trace into higher-level
architectural constructs, \eg, messages, operations, etc.: a
message such as {\tt Authorization request} may be
implemented in hardware through a Boolean or temporal
combination of specific hardware signals in the NoC fabric
between {\tt Device} and {\tt CE}, \eg, as a sequence
containing a header, a specific value of a sequence of data
words, etc.  We will refer to such architectural constructs
as {\em protocol events} or {\em flow events}.  Note that
due to limited observability, it may not be possible to map
a given set of (observed) hardware signals uniquely to a
flow event.  Finally, the trace may a result from several
instances of the same protocol executing concurrently, \eg,
a firmware authentication protocol may be invoked when
another instance of the protocol has not completed.

Trace interpretation entails mapping flow events created
during trace abstraction to system-level protocols in order
to identify the set of protocol instances (and
interleavings) responsible for creating the observed
behavior.  The trace may identify a problem in the protocols
themselves, \eg~an interleaving of some protocol executions
may lead to an unexpected message being sent or cause the
system to crash.  More commonly, one finds a bug in the {\em
  implementation} of the protocol, \ie, a trace inconsistent
with any possible interleaving of the protocol executions.
Identifying these problems involves significant human
expertise, and can often take days to weeks of effort.

@@

\noindent 
{\bf Notations and formalization.}  We assume that the set
of system flows is provided as a collection ${\vec{F}}$ of
LPNs.  We define a {\em flow execution scenario} to be a
set $\{(F_{i,j}, s_{i,j})\}$ where $F_{i,j}$ is the
$j$th instance of flow $F_i \in {\vec{F}}$, and $s_{i,j}$ is
a state of $F_{i,j}$.  A flow execution scenario indicates
the set of protocols and the number of instances of a
particular protocol are activated and their corresponding
current states.  Since we assume full observability, we view
an {\em observed trace} $\rho = e_1e_2\ldots e_n$ is a
sequence of events.  Given an observed trace $\rho$, the
goal of of trace interpretation is to construct a set of
candidate flow execution scenarios whose execution can
create the sequence of events in $\rho$.  We call such
execution scenarios {\em compliant} with $\rho$.  Let
$\mathit{accept(F_{i,j}, s_{i,j}, e)}$ be a function that
determines if event $e$ can be emitted by $F_{i,j}$ in state
$s_{i,j}$.  Formally, $\mathit{accept(F_{i,j}, s_{i,j}, e)}$
returns $(F_{i,j}, s'_{i,j})$ where $s'_{i,j} = (s_{i,j} - \bullet t)
\cup t\bullet$ if there exists a transition
$t$ in $F_i$ such that $L(t) = e$ and $\bullet t \subseteq
s_{i,j}$.  It returns $\emptyset$ if no such $t$
exists in $F_i$.

@@

Given an obsered trace $\rho$ and the set $\vec{F}$ of LPNs,
Algorithm~\ref{algo:compliance} provides a basic procedure for
computing a set of compliant flow execution scenarios.
\begin{algorithm}[ht]
\DontPrintSemicolon
Create an empty scenario $scen$\;
$\mathit{Scen} = \{scen\}$\;
\ForEach {$h, \; 1 \leq h \leq n $} {
	$\mathit{found} \gets {\tt true}$ \;
	$Scen' = \emptyset$\;
	\ForEach {$scen \in Scen$} {
  		\ForEach {$(F_{i,j}, s_{i,j}) \in \mathit{scen_1}$} {
    		\If {$\mathit{accept}(F_{i,j}, s_{i,j}, e_h) = (F_{i,j}, s'_{i,j})$} {
				Let $scen'$ be a copy of $scen$\;
      		$\mathit{scen'} \gets (\mathit{scen'} - (F_{i,j}, s_{i,j})) \cup (F_{i,j}, s'_{i,j})$\;
      		$\mathit{Scen'} \gets \mathit{scen'} \cup \mathit{Scen'}$\;
      		$\mathit{found} \gets {\tt false}$ \;
    		}
  		}
  		\ForEach {$F_i \in \vec{F}$} {
      	create a new instance $F_{i, j+1}$ \;
      	\If {$\mathit{accept}(F_{i,j+1},\mathit{init}_{i,j+1}, e_h) = (F_{i,j+1}, s'_{i,j+1})$} {
    			Let $scen'$ be a copy of $scen$\;
				$\mathit{scen'} \gets \mathit{scen'} \cup (F_{i,j+1}, s'_{i,j+1})$ \;
				$\mathit{Scen'} \gets \mathit{scen'} \cup \mathit{Scen'}$\;
        		$\mathit{found} \gets {\tt false}$ \;
      	}
    	}
	}
  \If {$\mathit{found} == {\tt true}$} {
    \Return ${\tt Inconsistent}$\;
  }
  $Scen = Scen'$\;
}
\Return $\mathit{scen}$ \;
\caption{$\textsc{Check-Compliance}(\vec{F}, \, \rho)$}
\label{algo:compliance}
\end{algorithm}
The algorithm operates by keeping track (in variable
$\mathit{Scen}$) of a set of candidate flow execution scenarios
compliant with each prefix of $\rho$.  At each iteration,
for each event $e_h$ in the observed trace, we
update $Scen$ by either extending a member of
$\mathit{scen}$ or initiating a new protocol instance for 
each $scen \in Scen$ with respect to $e_h$ in every possible way.
If $e_h$ cannot be emitted by any existing or new flow instances, 
then we report that the trace is {\tt  inconsistent}, \ie, 
there is no possible interleaving of
the protocol instances from $\vec{F}$ that is compliant with
$\rho$.

To illustrate the basic idea of the trace analysis method, 
consider the system flow shown in Figure~Fig.~\ref{flow-spec-ex}(b).
Let $F_1$ denote such flow.  Suppose that a hardware system 
implements flow $F_1$, and the following flow trace is abstracted 
from an observed signal trace as a result of executing such system.
\[
	t_1\;t_2\;t_1\;t_2\;t_3\;t_3\;t_4\;t_5\;t_5\;t_4\ldots
\]  
where transition names in the LPN are used to represent the 
flow events in the trace.  The first four events results in
the following flow execution scenario
\[
	\{(F_{1,1}, \{p_3\}),~(F_{1,2}, \{p_3\})\}.
\]
For the first event $t_3$, it results in two execution scenarios 
below depending on which flow instance emits $t_3$.
\[
\begin{array}{l}
	\{(F_{1,1}, \{p_4\}),~(F_{1,2}, \{p_3\})\} \\
	\{(F_{1,1}, \{p_3\}),~(F_{1,2}, \{p_4\})\}.
\end{array}
\]
After handing the next event $t_3$, the above two execution scenarios
are reduced to the one as shown below.
\[
	\{(F_{1,1}, \{p_4\}),~(F_{1,2}, \{p_4\})\}.
\]
By following Algorithm~\ref{algo:compliance} to handle the remaining 
four events, the following execution scenario can be derived.
\[
	\{(F_{1,1}, \{p_5, p_6\}),~(F_{1,2}, \{p_5, p_6\})\}
\]
%% Given a trace of flow events $\rho = e_1e_2\ldots e_n$, the
%% trace interpretation algorithm starts with an empty set of
%% of flow execution scenario $\mathit{Scen} = \emptyset$.
%% Then, for each $e_h$ where $1 \leq h \leq n$ starting $h=1$,
%% and for each $\mathit{scen} \in \mathit{Scen}$, the
%% following two steps are performed.
%% \begin{description}
%% \item[Step 1]~~For each $(F_{i,j}, s_{i,j}) \in
%%   \mathit{scen}$, if $\mathit{accept}(F_{i,j}, s_{i,j}, e_h)
%%   = (F_{i,j}, s'_{i,j})$, create a new scenario
%%   $\mathit{scen}' = (\mathit{scen} - (F_{i,j}, s_{i,j}))
%%   \cup (F_{i,j}, s'_{i,j})$, which is added into
%%   $\mathit{Scen}'$.

@@

%%-- Case study I
%\input{case1}

%
%--
%
\section{Case Study}
